---
title: "For the preparation of the paper submitted to Data in Brief"
date: "`r paste0('Last run at ', Sys.time(), ' on ', Sys.info()['nodename'], ', OS=',Sys.info()['sysname'])`"
output: 
  html_notebook:
    theme: united
    df_print: paged
    code_folding: hide
  toc: yes
---
```{r setup, include=FALSE}
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(tidyr))
suppressPackageStartupMessages(library(ggplot2))
```

# Data pre-processing
* Read the Raw data 'Raw.csv' under subfolder 'data'
```{r}
Raw_df <- read.csv(
          file='./data/Raw.csv',
          stringsAsFactors = T)
Raw_df$SessionID <- as.character(Raw_df$SessionID)
Raw_df$TimeStamp <- as.character(Raw_df$TimeStamp)
summary(Raw_df)
```
## Compute the impplied service time using the difference of the beginning time of two consecutive consultations in the same session
* Define the function: for a given session, use the consecutive time stamps to calculate service time
```{r}
calServTime <- function(session_df){
  stopifnot(length(unique(session_df$SessionID)) == 1) #stop if not the same session
  session_df <- session_df[order(session_df$TimeStamp),]
  TimeStamp = strptime(session_df$TimeStamp,"%H:%M:%S")
  session_df$ServTime<-c(as.numeric(difftime(TimeStamp[-1],
                                             TimeStamp[-nrow(session_df)],units = 'secs')),
                     NA)
  return(session_df[-nrow(session_df),])
}
```
* Compute the service time and add it as variable to the dataframe ServiceTime_df
```{r}
tmp_ID <- unique(Raw_df$SessionID)
ServiceTime_df <- data.frame()
for(ID in tmp_ID){
  ServiceTime_df<-rbind(ServiceTime_df,
                  calServTime(session_df = Raw_df %>% filter(SessionID ==ID))
  )
}
ServiceTime_df <- ServiceTime_df %>% select(-TimeStamp)
summary(ServiceTime_df)
```

* Save the data into the file 'ServiceTime.csv'
```{r}
write.csv(ServiceTime_df, 
          file='./data/ServiceTime.csv', row.names = F)
```

# Preliminary Analysis
## List of Tasks
1. Summary statistics of the 'number of consultations per session'
2. Number of sessions for each Day-Of-Week, and for Working Day vs Non-Working Day
3. Number of 'AM' sessions vs. that of 'PM' sessions
4. Number of consultation records grouped by 'Gender'
5. Number of consultation records grouped by 'M.Cancer' (M.Cancer=TRUE if and only if the visit was mainly for some sort of cancer)
6. Number of consultation records grouped by 'S.Cancer' (S.Cancer=TRUE if and only if the patient has some a certain type of cancer that was not the main reason for the consultation visit)
7. Number of consultation records grouped by 'Gender' and 'M.Cancer'
8. Number of consultation records grouped by 'Gender' and 'S.Cancer'
9. The distribution of 'Visit.No'. Explanation of 'Visit.No': by example. If a consultation record was the 3rd time this patient had been consulting this physician for the same reason (the same medical condition), then Visit.No=3 in this record.  
10. Service time distribution exploration: Overall distribution and the distribution faceting on various variables

## <span style="color:red"> When 'Service-Time' is not involved, we use all the records (from  Raw.csv </span>

## 1. Summary statistics of the 'number of consultations per session'
```{r}
Raw_df %>% 
  group_by(SessionID) %>%
  summarise(count = n()) %>% 
  summary()
```

## 2. Number of sessions for each Day-Of-Week
```{r}
Raw_df %>% 
  mutate(DayOfWeek=ordered(DayOfWeek,levels=c('Tuesday','Wednesday','Friday','Saturday'))) %>% 
  group_by(
  DayOfWeek
  ) %>% 
  summarise(num_Session = length(unique(SessionID)))
```

* 2.b Number of Working Day Sessions vs. the Non-Working Day one
```{r}
Raw_df %>% 
  mutate(WorkingDay=ordered(WorkingDay,levels=c('YES','NO'))) %>%
  group_by(
  WorkingDay
  ) %>% 
  summarise(num_Session = length(unique(SessionID)))
```

## 3. Number of 'AM' sessions vs. that of 'PM' sessions

```{r}
Raw_df %>% 
  mutate(AM_PM=ordered(AM_PM,levels=c('Morning','Afternoon'))) %>%
  group_by(
  AM_PM
  ) %>% 
  summarise(num_Session = length(unique(SessionID)))
```

## 4. Number of consultation records grouped by 'Gender'

```{r}
Raw_df %>% 
  select(Gender) %>% summary()
```

## 5. Number of consultation records grouped by 'M.Cancer' 
* Note M.Cancer=TRUE if and only if the visit was mainly for some sort of cancer

```{r}
Raw_df %>% select(
  M.Cancer
  ) %>% summary()
```

## 6. Number of consultation records grouped by 'S.Cancer' 
* Note: S.Cancer=TRUE if and only if the patient has some a certain type of cancer that was not the main reason for the consultation visit

```{r}
Raw_df %>% select(
  S.Cancer
  ) %>% summary()
```

# 7. Number of consultation records grouped by 'Gender' and 'M.Cancer'

```{r}
Raw_df %>% group_by(
  Gender,M.Cancer
  ) %>% tally() %>% 
  spread(
    Gender,
    n)
```

# 8. Number of consultation records grouped by 'Gender' and 'S.Cancer'

```{r}
Raw_df %>% group_by(
  Gender,S.Cancer
  ) %>% tally() %>% 
  spread(
    Gender,
    n)
```

# 9. The distribution of 'Visit.No'. 
* Explanation of 'Visit.No': by example. If a consultation record was the 3rd time this patient had been consulting this physician for the same reason (the same medical condition), then Visit.No=3 in this record.  
 
* Basic Summary of Visit.No
```{r}
Raw_df %>% select(Visit.No) %>% summary()
```
* Histogram of Visit.No
```{r}
old.par = par(no.readonly = T)
par(mfrow=c(1,2))
hist(Raw_df$Visit.No,main='Histogram of Visit.No', xlab='',col='blue')
boxplot(Raw_df$Visit.No,main='boxplot of Visit.No', xlab='',col='blue')
wanted_devices <- c("png", "pdf")
# Loop over all devices and copy the plot there
for (device in wanted_devices) {
  dev.copy(
    eval(parse(text = device)),
    paste("./figure/VisitNo", device, sep = ".")  # 
  )
  dev.off()
}
par(old.par)
```



# 10. Service time distribution exploration
* Note: since we are using the calculated service time data, we need to use ServiceTime_df instead of Raw_df.

* Overall distribution and the distribution faceting on various variables 
```{r}
ServiceTime_df %>% group_by(M.Cancer, Gender) %>%  summarise(count=n(),ServTime_Mean=mean(ServTime))
```

# ServiceTime Distribution faceting on M.Cancer
```{r,out.height="150%"}
old.par = par(no.readonly = T)
par(mfrow=c(2,1))
ServT_MainCancer = subset(ServiceTime_df$ServTime,ServiceTime_df$M.Cancer)
ServT_NotMainCancer = subset(ServiceTime_df$ServTime,!ServiceTime_df$M.Cancer)
xlim = range(c(ServT_MainCancer,ServT_NotMainCancer)) 
title_M_Cancer = sprintf('Service-Time (M.Cancer=TRUE): (mean,median)=(%.0f,%.0f) seconds',
                         mean(ServT_MainCancer),median(ServT_MainCancer))
title_notM_Cancer = sprintf('Service-Time (M.Cancer=FALSE): (mean,median)=(%.0f,%.0f) seconds',
                         mean(ServT_NotMainCancer),median(ServT_NotMainCancer))
hist(ServT_MainCancer,main=title_M_Cancer, 
     probability =TRUE, xlim= xlim,
     xlab='ServiceTime',col='blue')
hist(ServT_NotMainCancer,main=title_notM_Cancer, 
     probability =TRUE, xlim= xlim,
     xlab='ServiceTime',col='blue')
wanted_devices <- c("png", "pdf")
# Loop over all devices and copy the plot there
for (device in wanted_devices) {
  dev.copy(
    eval(parse(text = device)),
    paste("./figure/ServTime_M_Cancer", device, sep = ".")  # 
  )
  dev.off()
}
par(old.par)
```

# ServiceTime Distribution faceting on S.Cancer
```{r,out.height="150%"}
old.par = par(no.readonly = T)
par(mfrow=c(2,1))
ServT_SecondaryCancer = subset(ServiceTime_df$ServTime,ServiceTime_df$S.Cancer)
ServT_NotSecondaryCancer = subset(ServiceTime_df$ServTime,!ServiceTime_df$S.Cancer)
xlim = range(c(ServT_SecondaryCancer,ServT_NotSecondaryCancer)) 
title_M_Cancer = sprintf('Service-Time (S.Cancer=TRUE): (mean,median)=(%.0f,%.0f) seconds',
                         mean(ServT_SecondaryCancer),median(ServT_SecondaryCancer))
title_notM_Cancer = sprintf('Service-Time (S.Cancer=FALSE): (mean,median)=(%.0f,%.0f) seconds',
                         mean(ServT_NotSecondaryCancer),median(ServT_NotSecondaryCancer))
hist(ServT_SecondaryCancer,main=title_M_Cancer, 
     probability =TRUE, xlim= xlim,
     xlab='ServiceTime',col='blue')
hist(ServT_NotSecondaryCancer,main=title_notM_Cancer, 
     probability =TRUE, xlim= xlim,
     xlab='ServiceTime',col='blue')
wanted_devices <- c("png", "pdf")
# Loop over all devices and copy the plot there
for (device in wanted_devices) {
  dev.copy(
    eval(parse(text = device)),
    paste("./figure/ServTime_S_Cancer", device, sep = ".")  # 
  )
  dev.off()
}
par(old.par)
```

# ServiceTime Distribution faceting on 'Cancer': 
* M.Cancer and S.Cancer are considered together as 'Cancer'
```{r,out.height="150%"}
old.par = par(no.readonly = T)
par(mfrow=c(2,1))
Cancer_tag = ServiceTime_df$M.Cancer | ServiceTime_df$S.Cancer 
ServT_Cancer = subset(ServiceTime_df$ServTime,Cancer_tag)
ServT_NotCancer = subset(ServiceTime_df$ServTime,!Cancer_tag)
xlim = range(c(ServT_Cancer,ServT_NotCancer)) 
title_M_Cancer = sprintf('Service-Time (with Cancer): (mean,median)=(%.0f,%.0f) seconds',
                         mean(ServT_Cancer),median(ServT_Cancer))
title_notM_Cancer = sprintf('Service-Time (without Cancer): (mean,median)=(%.0f,%.0f) seconds',
                         mean(ServT_NotCancer),median(ServT_NotCancer))
hist(ServT_Cancer,main=title_M_Cancer, 
     probability =TRUE, xlim= xlim,
     xlab='ServiceTime',col='blue')
hist(ServT_NotCancer,main=title_notM_Cancer, 
     probability =TRUE, xlim= xlim,
     xlab='ServiceTime',col='blue')
wanted_devices <- c("png", "pdf")
# Loop over all devices and copy the plot there
for (device in wanted_devices) {
  dev.copy(
    eval(parse(text = device)),
    paste("./figure/ServTime_Cancer", device, sep = ".")  # 
  )
  dev.off()
}
par(old.par)
```

# The relationship between Service Time and Visit.No
```{r}
cor(ServiceTime_df %>% filter(!is.na(Visit.No)) %>% select(Visit.No,ServTime))
```

# Service Time: Visit.No = 1 vs Visit.No !=1
```{r}
ServiceTime_df %>% filter(!is.na(Visit.No)) %>%
  mutate(FirstVisit= Visit.No==1) %>% ggplot(aes(x=ServTime)) +
  geom_histogram(bins=30,color='black',fill='blue') + facet_grid(.~FirstVisit,labeller = label_both) +
  xlab('Service-Time') + theme_bw()
ggsave('./figure/ServTime_FirstVisit.pdf')
ggsave('./figure/ServTime_FirstVisit.png')
```

